Economic Evaluation of Mis-speficied Geographically Weighted Regression Models for Site-specific Nitrogen Management

### 2023-07-28 Writing/manuscript_gwr.rmd
final proofread and typo corrections

## Conceptual Framework
On-farm precision experimentation (OFPE) can provide detailed as-applied input and yield data at dense resolutions over entire fields, which if well analyzed may provide valuable information that can improve site-specific input application recommendations [@bullock2019data; @lacoste2021farm]. Consider a yield response function $y = f(\textbf{x}, \textbf{c})$, where following @bullock2000agronomic, $\textbf{x}$ is a vector of input management variables and $\textbf{c}$ is a vector of spatial variables describing field and soil characteristics^[For simplicity, we suppress discussion of the need to bring variance in weather data into yield response estimation.  Obviously, generation and use of weather data is necessary to provide meaningful probabilistic analysis and input management advice.]. Letting a field site $i$ be defined as a contiguous area characterized uniformly by a value $\textbf{c}_i$ of the vector of field and soil characteristics, its *site-specific yield response function* is the reduced form of the field's yield response function created when the value of the vector of characteristics variables is set at the site's value, $\textbf{c}_i$:  $f_i(\textbf{x}) \equiv f(\textbf{x}, \textbf{c}_i)$.  This conceptualization suggests two methods of estimating a field's site-specific yield responses. The first is to estimate a "map" of the field characteristics, for example thought of as a set of the field's $n \geq 1$ sites' values of the characteristics vector: $\{\textbf{c}_1,\dots, \textbf{c}_n\}$.  With that map, regressing yield on $\textbf{x}$ and $\textbf{c}$ generates an estimated "field-specific" yield response function $\hat{f}_(\textbf{x},\textbf{c})$. Estimates of site-specific yield response functions can then be obtained by inserting the value of the site's characteristic variables into the estimate of the field-specific response function: $\hat{f}_i(\textbf{x}) \equiv \hat{f}(\textbf{x},\textbf{c}_i)$. Alternatively, a site-specific yield response function might be estimated simply by taking (input rate, yield) data from points around a site of interest, and regressing yield on input application rates to attain the estimate $\tilde{f}_i (\textbf{x})$. Stated roughly, using GWR takes the latter approach, which has the advantage of not needing a costly map of field characteristics values.

@@

where $c_{j,i} = (\alpha_{j,i},\beta_{j,i},\gamma_{j,i},\tau_{j,i})$ is the vector of yield response function parameters. $\tau_{j,i}$ can be interpreted as the critical value of N rate at which cell $(j, i)$'s yield reaches its plateau, where yield no longer increases with N rate. $\varepsilon_{j, i}$ is a stochastic error term. The quadratic-plateau functional form implies a diminishing marginal product of N until a yield plateau, which are desirable properties for a yield response function [@frank1990comparison] and a pattern commonly observed in agronomic studies. Note that the cell-specific yield response function $f_{j, i} (N)$ is the reduced form of the meta response function $f$ when the characteristics vector takes on the value $c_{j, i}$. In other words, all the soil or field properties that could affect yield response to N are embedded in the simulated parameters.

@@

In each simulation, for each cell $(j, i)$, the simulation's true yield $y_{j, i}$ was generated based on the cell's assigned application rate $N_{j,i}$, true response parameters $\textbf{c}_{j,i} = (\alpha_{j,i}, \beta_{j,i}, \gamma_{j,i}, \tau_{j,i})$, and error term $\varepsilon_{j,i}$. This procedure generated a field-level, cell-specific data set, $\{(y_{j, i},N_{j, i} ):j=1,\dots,1440;k=1,\dots, 6 \}$. To maintain consistency with real-world OFPEs, it was assumed that harvesting equipment could only record yields accurately at an 18m $\times$ 12m (subplot) resolution. Since every cell in a plot received the same N application rate, then so did every cell within a subplot, and therefore $N_{j,1} = N_{j,2}= \dots = N_{j,6} = N_j$ for each subplot $j$. Letting $\mu_j$ denote the mean of the yields of the cells in subplot $j$, the analysis of each round's data was based on 1440 $(\mu_j,N_j )$ observations.

@@

$$\hat{N}^{*SCAM} \equiv argmax_N⁡ \big[p\cdot \hat{f}^{SCAM} (N)-w\cdot N\big]$$

@@

$$y_j= \alpha_j+ \beta N_j + \eta N_j^2+ \varepsilon_j ,$$

@@

To estimate the bias in the estimation of the economic value of GWR-based VRA, the true per-ha net revenues ("profits") of the SCAM-based URA and GWR-based VRA were calculated by substituting the estimated optimal nitrogen rates into the "true" yield response functions. Profits from the SCAM approach were

$$\Pi^{SCAM} = \frac{1}{8640} \sum_{j=1}^{1440} \sum_{k=1}^6 \big[p \cdot f_{j, i} (\hat{N}^{*SCAM})-w \cdot \hat{N}^{*SCAM}\big]$$

and profits from the GWR approach (either GWRR or GWRT) were,

$$\Pi^{GWR}=\frac{1}{8640} \sum_{j=1}^{1440} \sum_{k=1}^6 [p \cdot f_{j, i} (\hat{N}^{*GWR}_j)-w \cdot \hat{N}^{*GWR}_j]$$

@@

$$\hat{\Pi}^{SCAM} = \frac{1}{8640} \sum_{j=1}^{1440} \sum_{k=1}^6 \big[p \cdot \hat{f}^{GWR}_{j, i} (\hat{N}^{*SCAM} )-w \cdot \hat{N}^{*SCAM}\big]$$

$$\hat{\Pi}^{GWR}=\frac{1}{8640} \sum_{j=1}^{1440} \sum_{k=1}^6 [p \cdot \hat{f}^{GWR}_{j, i} (\hat{N}^{*GWR}_{j} )-w \cdot \hat{N}^{*GWR}_{j}]$$
where $\hat{f}^{GWR}_{j, i}(\cdot) = \hat{f}^{GWR}_j(\cdot)$ denotes the estimated yield response function by GWR (either GWRR or GWRT) at cell $(j, i)$ within subplot $j$.

The estimated value of the GWR-based VRA compared to the SCAM-based URA is defined as

$$\hat{\Delta}^{GWR} = \hat{\Pi}^{GWR} - \hat{\Pi}^{SCAM}$$

The bias in the estimation of the value of GWR-based VRA was calculated as follows:

$$bias = \hat{\Delta}^{GWR} - \Delta^{GWR}$$

@@

```{r, fig.id = "pi-dif-dist", fig.cap = "The value of GWR-based VRA over SCAM-based URA for GWRR and GWRT", out.width = "6.5in", fig.width = 6}
# knitr::include_graphics(here::here("GitControlled/Writing/Figures/g_value.png"))
g_value
```

```{r fig.id = "bias-est-eonr", fig.cap = "Average bias in the estimation of EONR by GWRR and GWRT", fig.width = 6}
# knitr::include_graphics(here::here("GitControlled/Writing/Figures/g_eonr_bias.png"))
g_eonr_bias
```

```{r fig.id = "true-vs-estimated-optn-gwr-r", fig.cap = "Comparison of Estimated and True EONR at the middle price ratio of 6.56", fig.width = 6}
# knitr::include_graphics(here::here("GitControlled/Writing/Figures/g_comp_eonr.png"))
g_comp_eonr
```

```{r fig.id = "bias-est-pi", fig.cap = "Bias in the estimation of the value of GWR-based VRA over SCAM-based URA for GWRR and GWRT", fig.width = 6}
# knitr::include_graphics(here::here("GitControlled/Writing/Figures/g_bias.png"))
g_bias
```

>>>>

## Conceptual Framework
On-farm precision experimentation (OFPE) can provide detailed as-applied input and yield data at dense resolutions over entire fields, which if well analyzed may provide valuable information that can improve site-specific input application recommendations [@bullock2019data; @lacoste2021farm]. Consider a yield response function $y = f(\textbf{x}, \textbf{c})$, where following @bullock2000agronomic, $\textbf{x}$ is a vector of input management variables and $\textbf{c}$ is a vector of spatial variables describing field and soil characteristics^[For simplicity, we suppress discussion of the need to bring variance in weather data into yield response estimation.  Obviously, generation and use of weather data is necessary to provide meaningful probabilistic analysis and input management advice.]. Letting a field site $i$ be defined as a contiguous area characterized uniformly by a value $\textbf{c}_i$ of the vector of field and soil characteristics, its *site-specific yield response function* is the reduced form of the field's yield response function created when the value of the vector of characteristics variables is set at the site's value, $\textbf{c}_i$:  $f_i(\textbf{x}) \equiv f(\textbf{x}, \textbf{c}_i)$.  This conceptualization suggests two methods of estimating a field's site-specific yield responses. The first is to estimate a "map" of the field characteristics, for example thought of as a set of the field's $n \geq 1$ sites' values of the characteristics vector: $\{\textbf{c}_1,\dots, \textbf{c}_n\}$.  With that map, regressing yield on $\textbf{x}$ and $\textbf{c}$ generates an estimated "field-specific" yield response function $\hat{f}(\textbf{x},\textbf{c})$. Estimates of site-specific yield response functions can then be obtained by inserting the value of the site's characteristic variables into the estimate of the field-specific response function: $\hat{f}_i(\textbf{x}) \equiv \hat{f}(\textbf{x},\textbf{c}_i)$. Alternatively, a site-specific yield response function might be estimated simply by taking (input rate, yield) data from points around a site of interest, and regressing yield on input application rates to attain the estimate $\hat{f}_i (\textbf{x})$. Stated roughly, using GWR takes the latter approach, which has the advantage of not needing a costly map of field characteristics values.

@@

where $\textbf{c}_{j,i} = (\alpha_{j,i},\beta_{j,i},\gamma_{j,i},\tau_{j,i})$ is the vector of yield response function parameters. $\tau_{j,i}$ can be interpreted as the critical value of N rate at which cell $(j, i)$'s yield reaches its plateau, where yield no longer increases with N rate. $\varepsilon_{j, i}$ is a stochastic error term. The quadratic-plateau functional form implies a diminishing marginal product of N until a yield plateau, which are desirable properties for a yield response function [@frank1990comparison] and a pattern commonly observed in agronomic studies. Note that the cell-specific yield response function $f_{j, i} (N)$ is the reduced form of the meta response function $f$ when the characteristics vector takes on the value $\textbf{c}_{j, i}$. In other words, all the soil or field properties that could affect yield response to N are embedded in the simulated parameters.

@@

In each simulation, for each cell $(j, i)$, the simulation's true yield $y_{j, i}$ was generated based on the cell's assigned application rate $N_{j,i}$, true response parameters $\textbf{c}_{j,i} = (\alpha_{j,i}, \beta_{j,i}, \gamma_{j,i}, \tau_{j,i})$, and error term $\varepsilon_{j,i}$. This procedure generated a field-level, cell-specific data set, $\{(y_{j, i},N_{j, i} ):j=1,\dots,1440;i=1,\dots, 6 \}$. To maintain consistency with real-world OFPEs, it was assumed that harvesting equipment could only record yields accurately at an 18m $\times$ 12m (subplot) resolution. Since every cell in a plot received the same N application rate, then so did every cell within a subplot, and therefore $N_{j,1} = N_{j,2}= \dots = N_{j,6} = N_j$ for each subplot $j$. Letting $\mu_j$ denote the mean of the yields of the cells in subplot $j$, the analysis of each round's data was based on 1440 $(\mu_j,N_j )$ observations.

@@

$$\hat{N}^{*SCAM} \equiv argmax_N⁡ \big[p\cdot \hat{f}^{SCAM} (N)-w\cdot N\big].$$

@@

$$y_j= \alpha_j+ \beta_j N_j + \eta N_j^2+ \varepsilon_j ,$$

@@

To estimate the bias in the estimation of the economic value of GWR-based VRA, the true per-ha net revenues ("profits") of the SCAM-based URA and GWR-based VRA were calculated by substituting the estimated optimal nitrogen rates into the "true" yield response functions. Profits from the SCAM approach were:

$$\Pi^{SCAM} = \frac{1}{8640} \sum_{j=1}^{1440} \sum_{i=1}^6 \big[p \cdot f_{j, i} (\hat{N}^{*SCAM})-w \cdot \hat{N}^{*SCAM}\big]$$

and profits from the GWR approach (either GWRR or GWRT) were:

$$\Pi^{GWR}=\frac{1}{8640} \sum_{j=1}^{1440} \sum_{i=1}^6 [p \cdot f_{j, i} (\hat{N}^{*GWR}_j)-w \cdot \hat{N}^{*GWR}_j]$$

@@

$$\hat{\Pi}^{SCAM} = \frac{1}{8640} \sum_{j=1}^{1440} \sum_{i=1}^6 \big[p \cdot \hat{f}^{GWR}_{j, i} (\hat{N}^{*SCAM} )-w \cdot \hat{N}^{*SCAM}\big]$$

$$\hat{\Pi}^{GWR}=\frac{1}{8640} \sum_{j=1}^{1440} \sum_{i=1}^6 [p \cdot \hat{f}^{GWR}_{j, i} (\hat{N}^{*GWR}_{j} )-w \cdot \hat{N}^{*GWR}_{j}]$$
where $\hat{f}^{GWR}_{j, i}(\cdot) = \hat{f}^{GWR}_j(\cdot)$ denotes the estimated yield response function by GWR (either GWRR or GWRT) at cell $(j, i)$ within subplot $j$.

The estimated value of the GWR-based VRA compared to the SCAM-based URA is defined as:

$$\hat{\Delta}^{GWR} = \hat{\Pi}^{GWR} - \hat{\Pi}^{SCAM}.$$

Finally, the bias in the estimation of the value of GWR-based VRA was calculated as follows:

$$bias = \hat{\Delta}^{GWR} - \Delta^{GWR}.$$

@@

```{r, fig.id = "pi-dif-dist", fig.cap = "The true economic value of GWR-based VRA over SCAM-based URA by two GWR models (GWRR and GWRT) and at three nitrogen-corn price ratios", out.width = "6.5in", fig.width = 6}
# knitr::include_graphics(here::here("GitControlled/Writing/Figures/g_value.png"))
g_value
```

```{r fig.id = "bias-est-eonr", fig.cap = "Average bias in the estimation of EONR by two GWR models (GWRR and GWRT) and at three nitrogen-corn price ratios", fig.width = 6}
# knitr::include_graphics(here::here("GitControlled/Writing/Figures/g_eonr_bias.png"))
g_eonr_bias
```

```{r fig.id = "true-vs-estimated-optn-gwr-r", fig.cap = "Comparison of GWR-Estimated and True EONR at the middle price ratio of 6.56", fig.width = 6}
# knitr::include_graphics(here::here("GitControlled/Writing/Figures/g_comp_eonr.png"))
g_comp_eonr
```

```{r fig.id = "bias-est-pi", fig.cap = "Bias in the estimation of the value of GWR-based VRA over SCAM-based URA by two GWR models (GWRR and GWRT) and at three nitrogen-corn price ratios", fig.width = 6}
# knitr::include_graphics(here::here("GitControlled/Writing/Figures/g_bias.png"))
g_bias
```
